Convolutional equivalent layer for gravity data processing

### 2019-12-03 manuscript/field_data.tex

\section{Field data test}
Test with real data are conducted with the gravity data from Caraj\'as, north of Brazil, were provided by the  Geological 
Survey of Brazil (CPRM). The real aerogravimetric data were collected in 113 flight lines along north–south direction 
with flight line spacing  of 3 km and tie lines along east–west direction at 12 km.

This airborne gravity survey was divided in two different areas, collected in different times, having samples spacing of 
$7.65$ m and $15.21$ m, totalizing  $4\,353\,428$ observation points. The height of the flight was fixed at 900 m. 
The gravity data (Figure \ref{fig:carajas_real_data}) were gridded into a regularly spaced dataset of $250\,000$ 
observation points ($500 \times 500$) with a grid spacing of $716.9311$ km north-south and $781.7387$ km east-west.

To apply our modification of the fast equivalent-layer method \cite[]{siqueira-etal2017}  that computes the forward 
modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}), we set an equivalent layer at 300 m deep. 
Figure \ref{fig:carajas_gz_predito_val}a shows the fitted gravity data after 50 iterations by applying our method. 
The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
(Figure \ref{fig:carajas_real_data}) and the predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show an 
acceptable data fitting because they have a mean close to zero ($0.0003$ mGal) and a small standard deviation of 
$0.105$ mGal which corresponds to approximately $0.1$ \% of the amplitude of the gravity data.

These small residuals indicate that our method yielded an estimated mass distribution (not shown) that can be used in 
the data processing. We perform upward-continuation of the real gravity data (Figure \ref{fig:carajas_real_data}) at 
a constant height of $5\,000$ m over the real data. The upward-continued gravity data 
(Figure \ref{fig:up2000_carajas_500x500}) seem a reasonable processing because of the attenuation of the short 
wavelenghts. By using our approach, the processing of the $250\,000$ observations was extremely fast and took 
$0.216$ seconds.

>>>>

\section{Application to field data}

We applied our method to airborne gravity data from Caraj{\'a}s, north of Brazil, which were provided by 
the Geological Survey of Brazil (CPRM). The data were collected along $113$ north–south flight lines separated 
by $3$ km and XXXXXX east–west tie lines separated by $12$ km.
This data set was divided in two different areas, collected in different times, having samples spacing of 
$7.65$ m and $15.21$ m along the lines, totalizing  $4\,353\,428$ observation points at a fixed height 
of $900$ m ($z_{1} = -900$ m). 
The gravity data were interpolated (Figure \ref{fig:carajas_real_data}) into a regularly spaced grid of 
$500 \times 500$ observation points ($N = 250\,000$) with a grid spacing of $716.9311$ m north-south and 
$781.7387$ m east-west.

To apply our method, we set an equivalent layer at $z_{0} = -300$ m. 
Figure \ref{fig:carajas_gz_predito_val}a shows the predicted data obtained with our method after 
$N^{it} = 50$ iterations.
The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
(Figure \ref{fig:carajas_real_data}) and predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show a 
very good data fit with mean close to zero ($0.0003$ mGal) and small standard deviation ($0.105$ mGal), 
which corresponds to approximately $0.1$ \% of the maximum amplitude of the gravity data.
By using the estimated mass distribution (not shown), we performed an upward-continuation of the 
observed gravity data to a horizontal plane located $5\,000$ m above. Figure \ref{fig:up2000_carajas_500x500} 
shows a very consistent upward-continued gravity data, with a clear attenuation of the short 
wave-lengths. By using our approach, the processing of the $250\,000$ observations took only 
$0.216$ seconds.

### 2019-12-03 manuscript/figures.tex

\plot{synthetic_data}{width=13.5cm}{
	{Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
	two rectangular prisms with density contrasts $0.35\, \mathrm{g/cm^3}$ (upper-left body) and 
	$0.4\, \mathrm{g/cm^3}$ (upper-right body). The black lines represent the horizontal projection 
	of the sources.}
	\label{fig:synthetic_data}
}

\plot{classic_fast_val}{width=8cm}{
	{(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
	respectively.}
	\label{fig:classic_fast_val}
}

\plot{bccb_fast_val}{width=8cm}{
	{(a) Predicted data produced by our method. 
	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
	respectively.}
	\label{fig:bccb_fast_val}
}

\plot{delta_comparison}{width=13.5cm}{
	{Difference between the predicted data produced by the fast equivalent-layer technique 
	(Figure \ref{fig:classic_fast_val}a) and by our method (Figure \ref{fig:bccb_fast_val}a).}
	\label{fig:delta_comparison}
}

\plot{delta_rho}{width=13.5cm}{
	{Difference between the estimated mass distribution within the equivalent layers produced by 
	the fast equivalent-layer technique (not shown) and by our method (not shown).}
	\label{fig:delta_rho}
}

\plot{upward_comparison}{width=5cm}{
	{The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
	panels a and b with mean $-5.938 \times 10^{-18}$ mGal and standard deviation $8.701 \times 10^{-18}$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and our method 
	are $7.62026$ and $0.00834$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_comparison}{width=5cm}{
	{The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
	panels a and b with mean $5.914 \times 10^{-18}$ mGal and standard deviation $9.014 \times 10^{-18}$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and our method 
	are $7.59654$ and $0.00547$ seconds, respectively.}
	\label{fig:downward_comparison}
}

@@

\plot{carajas_real_data}{width=13.5cm}{
	{Caraj\'as Province, Brazil. Gravity data on a regular grid of $500 \times 500$ points, totaling $250,000$ observations.
	The inset shows the study area (blue rectangle) which covers the southeast part of the state of Par\'a, north of Brazil.}
	\label{fig:carajas_real_data}
}

\plot{carajas_gz_predito_val}{width=8cm}{
	{Caraj\'as Province, Brazil. (a) Predicted gravity data produced by our modification of the fast equivalent-layer 
	method \citep{siqueira-etal2017}  that computes the forward modeling using the properties of BTTB and BCCB matrices
	(equation \ref{eq:w_Cv}). (b) Gravity residuals, defined as the difference between the observed data in 
	Figure \ref{fig:carajas_real_data} and the predicted data in panel a, with their mean of 0.000292 mGal and standard
	deviation of $0.105$ mGal.}
	\label{fig:carajas_gz_predito_val}
}

\plot{up2000_carajas_500x500}{width=13.5cm}{
	{Caraj\'as Province, Brazil. The upward-continued gravity data using our modification of the fast equivalent layer 
	method \citep{siqueira-etal2017} that computes the forward modeling using the properties of BTTB and BCCB matrices
	(equation \ref{eq:w_Cv}). The total computation time is 0.216 seconds for processing of the $250,000$ observations.}
	\label{fig:up2000_carajas_500x500}
}

>>>>

\plot{synthetic_data}{width=13.5cm}{
	{Application to synthetic data. 
	Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
	two rectangular prisms with density contrasts $0.35\, \mathrm{g/cm^3}$ (upper-left body) and 
	$0.4\, \mathrm{g/cm^3}$ (upper-right body). The black lines represent the horizontal projection 
	of the sources.}
	\label{fig:synthetic_data}
}

\plot{classic_fast_val}{width=8cm}{
	{Application to synthetic data. 
	(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
	respectively.}
	\label{fig:classic_fast_val}
}

\plot{bccb_fast_val}{width=8cm}{
	{Application to synthetic data. 
	(a) Predicted data produced by our method. 
	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
	respectively.}
	\label{fig:bccb_fast_val}
}

\plot{delta_comparison}{width=13.5cm}{
	{Application to synthetic data. 
	Difference between the predicted data produced by the fast equivalent-layer technique 
	(Figure \ref{fig:classic_fast_val}a) and by our method (Figure \ref{fig:bccb_fast_val}a).}
	\label{fig:delta_comparison}
}

\plot{delta_rho}{width=13.5cm}{
	{Application to synthetic data. 
	Difference between the estimated mass distribution within the equivalent layers produced by 
	the fast equivalent-layer technique (not shown) and by our method (not shown).}
	\label{fig:delta_rho}
}

\plot{upward_comparison}{width=5cm}{
	{Application to synthetic data. 
	The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
	panels a and b with mean $-5.938 \times 10^{-18}$ mGal and standard deviation $8.701 \times 10^{-18}$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and our method 
	are $7.62026$ and $0.00834$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_comparison}{width=5cm}{
	{Application to synthetic data. 
	The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
	panels a and b with mean $5.914 \times 10^{-18}$ mGal and standard deviation $9.014 \times 10^{-18}$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and our method 
	are $7.59654$ and $0.00547$ seconds, respectively.}
	\label{fig:downward_comparison}
}

@@

\plot{carajas_real_data}{width=13.5cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	Observed gravity data on a regular grid of $500 \times 500$ points, totaling $N = 250,000$ 
	observations. The inset shows the study area (blue rectangle) which covers the southeast part of 
	the state of Par{\'a}, north of Brazil.}
	\label{fig:carajas_real_data}
}

\plot{carajas_gz_predito_val}{width=8cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	(a) Predicted data produced by our method. (b) Residuals between the observed  
	(Figure \ref{fig:carajas_real_data}) and the predicted data (panel a), with mean 
	$0.000292$ mGal and standard deviation of $0.105$ mGal.}
	\label{fig:carajas_gz_predito_val}
}

\plot{up2000_carajas_500x500}{width=13.5cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	The upward-continued gravity data obtained with our method $5\,000$ m above the 
	observed data (Figure \ref{fig:carajas_real_data}).
	The total computation time for processing of the $250,000$ observations was $0.216$ seconds.}
	\label{fig:up2000_carajas_500x500}
}
